### 2024-02-06 docs/projects/noj/ggplot.qmd
Noj - rerendered

::: {.sourceClojure}
```clojure
(ns ggplot
  (:require [clojisr.v1.r :as r :refer [r r$ r->clj]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.walk :as walk]
            [tablecloth.api :as tc]))
```
:::

@@

<g id="surface931">

@@

::: {.sourceClojure}
```clojure
(defn ggolot->clj
  ([r-obj
    options]
   (ggolot->clj r-obj options []))
  ([r-obj
    {:as options
     :keys [avoid]}
    path]
   #_(prn path)
   (let [relevant-names (some->> r-obj
                                 r.base/names
                                 r->clj
                                 (filter (complement avoid)))]
     (cond
       ;;
       ;; a named list or a ggproto object
       (seq relevant-names) (->> relevant-names
                                 (map (fn [nam]
                                        [(keyword nam) (-> r-obj
                                                           (r$ nam)
                                                           (ggolot->clj options
                                                                        (conj path nam)))]))
                                 (into {}))
       ;;
       ;; a ggproto method
       (-> r-obj
           r.base/class
           r->clj
           first
           (= "ggproto_method"))
       :ggproto-method
       ;;
       ;; an unnamed list
       (-> r-obj
           r.base/is-list
           r->clj
           first)
       (-> r-obj
           r.base/length
           r->clj
           first
           range
           (->> (mapv (fn [i]
                        (prn [path (inc i)])
                        (-> r-obj
                            (r/brabra (inc i))
                            (ggolot->clj options
                                         (conj path [i])))))))
       ;;
       (r.base/is-atomic r-obj) (try (r->clj r-obj)
                                     (catch Exception e
                                       (-> r-obj println with-out-str)))
       :else r-obj))))
```
:::

@@

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
    r
    (ggolot->clj {:avoid #{"plot_env"}})
    (dissoc :data))
```
:::

@@

::: {.sourceClojure}
```clojure
(defn h4 [title]
  (kind/hiccup [:h3 title]))
```
:::

::: {.sourceClojure}
```clojure
(defn ggplot-summary [r-code]
  (let [plot (r r-code)
        clj (-> plot
                (ggolot->clj {:avoid #{"plot_env"}})
                (dissoc :data))]
    (kind/fragment
     [(h4 "R code")
      (kind/md
       (format "\n```\n%s\n```\n"
               r-code))
      (h4 "plot")
      (-> plot
          plotting/plot->buffered-image)
      (h4 "clj data")
      (kind/pprint clj)])))
```
:::

### A scatterplot

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy))
         + geom_point())")
```
:::

```{=html}
<h3>R code</h3>
```

@@

```{=html}
<h3>plot</h3>
```

@@

```{=html}
<h3>clj data</h3>
```

@@

### A scatterplot with colours

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())")
```
:::

```{=html}
<h3>R code</h3>
```

@@

### A scatterplot with colours and smoothing

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\"))")
```
:::

```{=html}
<h3>R code</h3>
```

@@

### A scatterplot with colours, smoothing, and facets

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))")
```
:::

```{=html}
<h3>R code</h3>
```

@@

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode: 0x559b84332b20&gt;\n&lt;environment: namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

>>>>

::: {.sourceClojure}
```clojure
(ns ggplot
  (:require [clojisr.v1.r :as r :refer [r r$ r->clj]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.walk :as walk]
            [tablecloth.api :as tc]
            [editscript.core :as editscript]
            [clojure.pprint :as pp]
            [clojure.string :as str]))
```
:::

@@

<g id="surface3322">

@@

::: {.sourceClojure}
```clojure
(defn ggplot->clj
  ([r-obj
    options]
   (ggplot->clj r-obj options []))
  ([r-obj
    {:as options
     :keys [avoid]}
    path]
   #_(prn path)
   (let [relevant-names (some->> r-obj
                                 r.base/names
                                 r->clj
                                 (filter (complement avoid)))]
     (cond
       ;;
       ;; a named list or a ggproto object
       (seq relevant-names) (->> relevant-names
                                 (map (fn [nam]
                                        [(keyword nam) (-> r-obj
                                                           (r$ nam)
                                                           (ggplot->clj options
                                                                        (conj path nam)))]))
                                 (into {}))
       ;;
       ;; a ggproto method
       (-> r-obj
           r.base/class
           r->clj
           first
           (= "ggproto_method"))
       :ggproto-method
       ;;
       ;; an unnamed list
       (-> r-obj
           r.base/is-list
           r->clj
           first)
       (-> r-obj
           r.base/length
           r->clj
           first
           range
           (->> (mapv (fn [i]
                        (-> r-obj
                            (r/brabra (inc i))
                            (ggplot->clj options
                                         (conj path [i])))))))
       ;;
       (r.base/is-atomic r-obj) (try (r->clj r-obj)
                                     (catch Exception e
                                       (-> r-obj println with-out-str)))
       :else r-obj))))
```
:::

@@

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
    r
    (ggplot->clj {:avoid #{"plot_env"}})
    (dissoc :data))
```
:::

@@

::: {.sourceClojure}
```clojure
(defn h3 [title] (kind/hiccup [:h3 title]))
```
:::

::: {.sourceClojure}
```clojure
(defn h4 [title] (kind/hiccup [:h4 title]))
```
:::

::: {.sourceClojure}
```clojure
(defn ggplot-summary
  ([title r-code]
   (ggplot-summary r-code))
  ([title r-code prev-clj-to-compare]
   (let [plot (r r-code)
         clj (-> plot
                 (ggplot->clj {:avoid #{"plot_env"}})
                 (dissoc :data))]
     {:title title
      :r-code r-code
      :image (plotting/plot->buffered-image plot)
      :clj clj
      :diff (when prev-clj-to-compare
              (-> prev-clj-to-compare
                  (editscript/diff clj)
                  pp/pprint
                  with-out-str
                  (str/replace #": " ":_ ")
                  read-string))})))
```
:::

::: {.sourceClojure}
```clojure
(defn view-summary [{:keys [title r-code image clj diff]}]
  (kind/fragment
   [(h3 title)
    (h4 "R code")
    (kind/md
     (format "\n```\n%s\n```\n"
             r-code))
    (h4 "plot")
    image
    (h4 "clj data")
    clj
    (when diff
      (kind/fragment
       [(h4 "clj diff with previous")
        diff]))]))
```
:::

::: {.sourceClojure}
```clojure
(->> [;;
      ["A scatterplot"
       "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"]
      ;;
      ["A scatterplot with colours"
       "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())"]
      ;;
      ["A scatterplot with colours and smoothing"
       "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\"))"]
      ;;
      ["A scatterplot with colours, smoothing, and facets"
       "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))"]]
     (reductions (fn [prev-summary [title r-code]]
                   (ggplot-summary title
                                   r-code
                                   (:clj prev-summary)))
                 nil)
     rest
     (map view-summary)
     kind/fragment)
```
:::

```{=html}
<h3>A scatterplot</h3>
```

```{=html}
<h4>R code</h4>
```

@@

```{=html}
<h4>plot</h4>
```

@@

```{=html}
<h4>clj data</h4>
```

@@

```{=html}
<h3>A scatterplot with colours</h3>
```

```{=html}
<h4>R code</h4>
```

@@

```{=html}
<h4>clj diff with previous</h4>
```

::: {.printedClojure}
```clojure
[[[:mapping :colour] :+ [~[factor cyl]]]
 [[:labels :colour] :+ ["factor(cyl)"]]]

```{=html}
<h3>A scatterplot with colours and smoothing</h3>
```

```{=html}
<h4>R code</h4>
```

@@

```{=html}
<h4>clj diff with previous</h4>
```

::: {.printedClojure}
```clojure
[[[:layers 1]
  :+
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {:_ nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}]]

```{=html}
<h3>A scatterplot with colours, smoothing, and facets</h3>
```

```{=html}
<h4>R code</h4>
```

@@

```{=html}
<h4>clj data</h4>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode: 0x557f413a3978&gt;\n&lt;environment: namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

```{=html}
<h4>clj diff with previous</h4>
```

::: {.printedClojure}
```clojure
[[[:facet :params]
  :+
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode:_ 0x557f413a3978&gt;\n&lt;environment:_ namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"}]]

### 2024-02-06 docs/projects/noj/ml.qmd
Noj - rerendered

Residuals:
       Min          1Q      Median          3Q         Max
   -0.7855     -0.2024      0.0281      0.2007      0.6159

Coefficients:
                  Estimate Std. Error    t value   Pr(>|t|)
Intercept           1.8316     0.3045     6.0143     0.0000 ***
sepal-width         0.6567     0.0798     8.2255     0.0000 ***
petal-length        0.6798     0.0680     9.9990     0.0000 ***
petal-width        -0.4501     0.1530    -2.9415     0.0041 **
---------------------------------------------------------------------
Significance codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 0.2950 on 96 degrees of freedom
Multiple R-squared: 0.8767,    Adjusted R-squared: 0.8728
F-statistic: 227.5000 on 4 and 96 DF,  p-value: 1.726e-43

>>>>

Residuals:
       Min          1Q      Median          3Q         Max
   -0.8076     -0.2157      0.0177      0.2335      0.7341

Coefficients:
                  Estimate Std. Error    t value   Pr(>|t|)
Intercept           1.8682     0.2970     6.2904     0.0000 ***
sepal-width         0.6388     0.0786     8.1279     0.0000 ***
petal-length        0.7250     0.0713    10.1635     0.0000 ***
petal-width        -0.5999     0.1646    -3.6449     0.0004 ***
---------------------------------------------------------------------
Significance codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 0.3197 on 96 degrees of freedom
Multiple R-squared: 0.8499,    Adjusted R-squared: 0.8452
F-statistic: 181.1746 on 4 and 96 DF,  p-value: 2.137e-39
